---
title: "Efficiency Regression"
output: html_document
---

```{r}
library(tidyverse)
library(fixest)
library(stargazer)
options(dplyr.summarise.inform = FALSE)
```

Regression table functions
```{r}
etable_iid <- function(models){
  etable(models,
         # cluster = cluster,
         # dict = dict_names,
         order = c("Age*", "!Intercept"),
         digits = "r3",
         digits.stats = 2,
         fitstat = c("n", "ar2", "war2"),
         style.tex = style.tex("aer",
                               fixef.suffix = " FEs",
                               fixef.where = "var",
                               yesNo = c("Yes", "No")),
         tex = F) %>%
      print()
}

etable_cluster <- function(models, cluster){
  etable(models,
         cluster = cluster,
         # dict = dict_names,
         order = c("Age*", "!Intercept"),
         digits = "r3",
         digits.stats = 2,
         fitstat = c("n", "ar2", "war2"),
         style.tex = style.tex("aer",
                               fixef.suffix = " FEs",
                               fixef.where = "var",
                               yesNo = c("Yes", "No")),
         tex = F) %>%
      print()
}
```

```{r}
load("data/All_join.Rda")
```

# Pivot and calculate
```{r}
error_criterion = 10

id_cols <- c("Fleet.Type",
             "CVN",
             "IMO.Number",
             "MRV.IMO.Number",
             "Name",
             "Flag.State",
             "MMSI",
             "EIV.Calculated",
             "Main.Engine.Power.kW",
             "Built.Year",
             "Size.Category",
             "Dwt",
             "Service.Speed..knots.",
             "LDT",
             "Draught..m.",
             "LOA..m.", # length overall
             "LBP..m.", # length between perpendiculars
             "Beam.Mld..m.",
             "Depth.Moulded..m.", # distance from the keel to the freeboard deck measured at the ship side
             "Beam.Ext..m.",
             "TPC", # tonnage per centimeter
             "NT", # net tonnage: dimensionless index calculated from the total moulded volume of the ship's cargo spaces
             "Main.Engine.Fuel.Type",
             "Main.Consumption.at.Laden.Speed..tpd.",
             "Main.Consumption.at.Ballast.Speed..tpd.",
             "Main.Consumption.at.Eco.Speed.Ballast..tpd.",
             "Main.Consumption.at.Eco.Speed.Laden..tpd.",
             "Main.Consumption.at.Service.Speed..tpd..",
             "Consumption..tpd.",
             "EST.Number..Energy.Saving.Technologies.",
             "Eco...Electronic.Engine",
             "P.and.I.Club.")

AIS_stats_df <- fleet_df %>% 
    filter(Fleet.Type == "Bulkers") %>% 
    drop_na(MMSI) %>%
    select(all_of(id_cols),
           contains("AIS"),
           contains("MRV.EU"),
           contains("MRV.Total.FC"),
           contains("MRV.Load"),
           -AIS.MMSI,
           -contains("AIS.Timestamp")) %>%
  pivot_longer(cols = -all_of(id_cols),
               names_pattern = "(.*)\\.([0-9]*)",
               names_to = c("Variable", "Year")) %>% 
  pivot_wider(names_from = c("Variable"),
              values_from = "value") %>% 
  mutate(EU.Distance.Error = AIS.EU.Distance - MRV.EU.Distance,
         EU.Distance.PctError = (AIS.EU.Distance/MRV.EU.Distance - 1) * 100,
         FC.Per.Travel.Work = MRV.Total.FC/AIS.EU.Work,
         Age = as.numeric(Year) - Built.Year,
         Electronic.Engine = Eco...Electronic.Engine == 
           "Eco – Electronic Engine" | Eco...Electronic.Engine == "Eco – Electronic Engine Modern",
         Average.Load = MRV.Load/AIS.EU.Trips) %>% 
  drop_na(AIS.EU.Distance, MRV.EU.Distance) %>% 
  mutate(Best = abs(EU.Distance.PctError) <= error_criterion)
  # mutate(Best = between(EU.Distance.PctError, -10, 5))
```

# Verify data
```{r}
AIS_stats_df %>% 
  ggplot(aes(EU.Distance.PctError,
             fill = Best)) +
  geom_histogram(position = 'identity',
                 alpha = 0.5,
                 binwidth = 1) +
  facet_wrap(~Year) +
  xlim(-200, 200)
```
```{r}
AIS_stats_df %>% 
  group_by(Year, Best) %>% 
  summarise(n())
```


# Data selection
```{r}
AIS_stats_best_df <- AIS_stats_df %>% 
  filter(Best) %>% 
  select(MMSI,
         Year,
         EU.Distance.PctError,
         AIS.EU.Work,
         MRV.Total.FC,
         FC.Per.Travel.Work,
         EIV.Calculated,
         Main.Engine.Power.kW,
         Built.Year,
         Size.Category,
         Age,
         Dwt,
         Service.Speed..knots.,
         MRV.Load,
         LDT,
         Draught..m.,
         LOA..m., # length overall
         LBP..m.,
         Beam.Mld..m.,
         Depth.Moulded..m.,
         # Beam.Ext..m., # Too few observations
         TPC, # tonnage per centimeter
         NT, # net tonnage: dimensionless index calculated from the total moulded volume of the ship's cargo spaces
         Main.Engine.Fuel.Type,
         Main.Consumption.at.Laden.Speed..tpd.,
         Main.Consumption.at.Ballast.Speed..tpd.,
         # Main.Consumption.at.Eco.Speed.Ballast..tpd., # Too few observations
         # Main.Consumption.at.Eco.Speed.Laden..tpd., # Too few observations
         Main.Consumption.at.Service.Speed..tpd..,
         Consumption..tpd.,
         EST.Number..Energy.Saving.Technologies.,
         Electronic.Engine,
         Average.Load,
         # AIS.Is.Gt24,
         P.and.I.Club.)
```

## Filter Outliers
```{r}
AIS_stats_best_df <- AIS_stats_best_df %>% 
  filter(between(FC.Per.Travel.Work,
                 quantile(AIS_stats_best_df$FC.Per.Travel.Work, probs = .01),
                 quantile(AIS_stats_best_df$FC.Per.Travel.Work, probs = .99)))
```

```{r}
AIS_stats_df %>% 
  mutate(Data = 'All') %>% 
  bind_rows(AIS_stats_best_df %>% mutate(Data = 'Best')) %>%
  ggplot(aes(EU.Distance.PctError,
             fill = Data)) +
  geom_histogram(binwidth = 1,
                 position = 'identity',
                 alpha = 0.6) +
  facet_wrap(~Year) +
  xlim(-200, 200)
```
Fraction of well-matched observations
```{r}
print(paste0(
  round(AIS_stats_df %>% filter(Best) %>% nrow() / AIS_stats_df %>% nrow(), 3),
  ' of ',
  AIS_stats_df %>% nrow(),
  ' ship-years have distance within +/-',
  error_criterion,
  '%'))
```


Number of ships with well-matched observations in both years
```{r}
AIS_stats_best_df %>%
  group_by(MMSI) %>% 
  summarise(count = n()) %>% 
  group_by(count) %>% 
  summarise(n())
```

# Derived variables
Block coefficient = $\frac{Displacment~Volume}{LBP*B*Draught}$
where $Displacement~Volume = Dwt + LDT$
Wetted area: $A_s = 1.025 LBP (C_B B + 1.7 Draught)$
```{r}
AIS_stats_best_df <- AIS_stats_best_df %>% 
  mutate(Block.Coeff.Dwt = Dwt/(LBP..m.*Beam.Mld..m.*Draught..m.),
         Block.Coeff.Dwt.LDT = (Dwt + LDT)/(LBP..m.*Beam.Mld..m.*Draught..m.),
         Block.Coeff.NT.LDT = (NT + LDT)/(LBP..m.*Beam.Mld..m.*Draught..m.),
         Wetted.Surface.Dwt = LBP..m.*(Block.Coeff.Dwt*Beam.Mld..m. + 1.7*Draught..m.),
         Wetted.Surface.Dwt.LDT = LBP..m.*(Block.Coeff.Dwt.LDT*Beam.Mld..m. + 1.7*Draught..m.))
```

# Plots
```{r}
AIS_stats_best_df %>% 
  ggplot(aes(AIS.EU.Work, fill = Year)) +
  geom_histogram(binwidth = 1E6) +
  # xlim(0, 1E7) +
  facet_wrap(~Size.Category, scales = 'free')
```

```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = MRV.Total.FC, fill = Year)) +
  geom_histogram(binwidth = 100) + 
  facet_wrap(~Size.Category)
```

```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = FC.Per.Travel.Work, fill = Year)) +
  geom_histogram() + 
  facet_wrap(~Size.Category)
  # xlim(0, 0)
```

```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = EIV.Calculated,
             y = FC.Per.Travel.Work,
             color = Year)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~Size.Category)
```

```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = Built.Year,
             y = FC.Per.Travel.Work,
             color = Year)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~Size.Category)
```
```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = Age,
             y = FC.Per.Travel.Work,
             color = Year)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~Size.Category)
```
```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = Dwt,
             y = FC.Per.Travel.Work,
             color = Size.Category)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~Year)
```

```{r}
AIS_stats_best_df %>% 
  distinct(MMSI, .keep_all = TRUE) %>% 
  ggplot(aes(y = Service.Speed..knots.,
             x = Main.Engine.Power.kW,
             color = Size.Category)) +
  geom_point() +
  geom_smooth(method = 'lm')
```
### Consumption v Built
```{r}
AIS_stats_best_df %>% 
  distinct(MMSI, .keep_all = TRUE) %>% 
  ggplot(aes(x = Built.Year, y = Consumption..tpd., color = Size.Category)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

### Block coeff v Built
```{r}
AIS_stats_best_df %>% 
  distinct(MMSI, .keep_all = TRUE) %>% 
  mutate(Block.Coeff = (Dwt + LDT)/(LBP..m.*Beam.Mld..m.*Draught..m.)) %>% 
  ggplot(aes(x = Built.Year, y = Block.Coeff, color = Size.Category)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
AIS_stats_best_df %>% 
  distinct(MMSI, .keep_all = TRUE) %>% 
  mutate(Block.Coeff = (Dwt + LDT)/(LBP..m.*Beam.Mld..m.*Draught..m.)) %>% 
  ggplot(aes(y = Consumption..tpd., x = Block.Coeff, color = Size.Category)) +
  geom_point() +
  geom_smooth(method = 'lm')
```
```{r}
AIS_stats_best_df %>% 
  distinct(MMSI, .keep_all = TRUE) %>% 
  ggplot(aes(x = Dwt, y = Dwt/LDT)) +
  geom_point()
```
```{r}
AIS_stats_best_df %>% 
  distinct(MMSI, .keep_all = TRUE) %>% 
  ggplot(aes(y = LDT/Dwt, x = Built.Year)) +
  geom_point() +
  geom_smooth(method = 'lm')
```
## Average Load
```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = Average.Load, fill = Year)) +
  geom_histogram(position = 'identity',
                 alpha = 0.4) +
  facet_wrap(~Size.Category, scales = 'free_x')
```
```{r}
AIS_stats_best_df %>% 
  ggplot(aes(x = Average.Load/Dwt*1000, fill = Year)) +
  geom_histogram(position = 'identity',
                 alpha = 0.4) +
  facet_wrap(~Size.Category)
  # xlim(0, 1)
```

## Histograms of regression variables
```{r}
AIS_stats_best_df %>% 
  select(-c(Size.Category, AIS.EU.Work, EU.Distance.PctError, Main.Engine.Fuel.Type, P.and.I.Club.)) %>%
  # filter(EIV.Calculated < 20,
  #        MRV.Total.FC < 9000,
  #        MRV.Load < 300) %>% 
  pivot_longer(-c(MMSI, Year)) %>% 
  ggplot(aes(value, fill = Year)) +
  geom_histogram(position = 'identity',
                 alpha = 0.5) +
  facet_wrap(~name, scales = 'free')
```
```{r}
AIS_stats_best_df %>% 
  select(MMSI, Year, MRV.Total.FC) %>%
  # filter(EIV.Calculated < 20,
  #        MRV.Total.FC < 9000,
  #        MRV.Load < 300) %>% 
  pivot_longer(-c(MMSI, Year)) %>% 
  ggplot(aes(value, fill = Year)) +
  geom_histogram(position = 'identity',
                 alpha = 0.5) +
  facet_wrap(~name, scales = 'free')
```

```{r}
odd_mmsi <- AIS_stats_best_df %>% 
  filter(EIV.Calculated > 20) %>% 
  pull(MMSI)

AIS_stats_df %>% 
  select(MMSI, MRV.EU.Distance, Main.Engine.Power.kW, Dwt, Service.Speed..knots., EIV.Calculated) %>% 
  filter(MMSI %in% odd_mmsi) %>% 
  arrange(desc(EIV.Calculated))
```


```{r}
odd_mmsi <- AIS_stats_best_df %>% 
  filter(MRV.Total.FC > 9000) %>% 
  pull(MMSI)

AIS_stats_df %>% 
  select(MMSI, MRV.EU.Distance, Dwt, MRV.Total.FC) %>% 
  filter(MMSI %in% odd_mmsi) %>% 
  arrange(desc(MRV.Total.FC))
```


```{r}
odd_mmsi <- AIS_stats_best_df %>% 
  filter(MRV.Load > 200) %>% 
  pull(MMSI)

AIS_stats_df %>% 
  select(MMSI, MRV.EU.Distance, Dwt, MRV.Load) %>% 
  filter(MMSI %in% odd_mmsi) %>% 
  arrange(desc(MRV.Load))
```
```{r}
odd_mmsi <- AIS_stats_best_df %>% 
  filter(FC.Per.Travel.Work > 0.002) %>% 
  pull(MMSI)

AIS_stats_df %>% 
  select(MMSI, MRV.EU.Distance, Dwt, FC.Per.Travel.Work) %>% 
  filter(MMSI %in% odd_mmsi) %>% 
  arrange(desc(FC.Per.Travel.Work))
```

# Regressions
# Preferred Specs
```{r}
models_pref <- feols(log(FC.Per.Travel.Work) ~ 
            csw(Age +
                log(Main.Engine.Power.kW) +
                log(Dwt) +
                log(LBP..m.) +
                log(Beam.Mld..m.) +
                log(Draught..m.) +
                log(MRV.Load),
                # log(Average.Load),
                log(TPC),
                log(Service.Speed..knots.),
                log(Consumption..tpd.))
            | sw0(Size.Category),
               data = AIS_stats_best_df)
etable_cluster(models_pref, 'MMSI')
```

# Most preferred?
```{r}
model_pref <- feols(log(FC.Per.Travel.Work) ~ 
                      Age +
                      log(Dwt) +
                      log(Main.Engine.Power.kW) +
                      log(LBP..m.) +
                      log(Beam.Mld..m.) +
                      log(Draught..m.) +
                      log(MRV.Load) +
                      log(TPC)
                      | Size.Category,
               data = AIS_stats_best_df %>% drop_na(TPC))
etable_iid(model_pref)
```
## Clustered Errors
```{r}
etable_cluster(model_pref, 'MMSI')
etable_cluster(model_pref, 'Size.Category')
```

```{r}
plot(fixef(model_pref))
```
## Analysis plots
```{r}
model_pref <- lm(log(FC.Per.Travel.Work) ~ 
                      Age +
                      log(Main.Engine.Power.kW) +
                      log(Dwt) +
                      log(LBP..m.) +
                      log(Beam.Mld..m.) +
                      log(Draught..m.) +
                      log(MRV.Load) + 
                      log(TPC) +
                      Size.Category,
               data = AIS_stats_best_df)
plot(model_pref)
d<-density(model_pref[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')
```

# Built Year Cohorts
5 year
```{r}
AIS_stats_best_df$Built.Cohort = cut(AIS_stats_best_df$Built.Year,
                                     breaks = seq(min(AIS_stats_best_df$Built.Year), 2021, by = 5))

models <- feols(log(FC.Per.Travel.Work) ~ 
                  csw(Age +
                      Built.Cohort +
                      log(Main.Engine.Power.kW) +
                      log(Dwt) +
                      log(LBP..m.) +
                      log(Beam.Mld..m.) +
                      log(Draught..m.) +
                      log(MRV.Load),
                      # log(Average.Load),
                      log(TPC),
                      log(Service.Speed..knots.),
                      log(Consumption..tpd.))
                  | Size.Category,
               data = AIS_stats_best_df)
etable_iid(models)
etable_cluster(models, 'Built.Cohort')
```
```{r}
AIS_stats_best_df$Built.Cohort = cut(AIS_stats_best_df$Built.Year,
                                     breaks = seq(min(AIS_stats_best_df$Built.Year), 2021, by = 9))

models <- feols(log(FC.Per.Travel.Work) ~ 
                  csw(Age +
                      Built.Cohort +
                      log(Main.Engine.Power.kW) +
                      log(Dwt) +
                      log(LBP..m.) +
                      log(Beam.Mld..m.) +
                      log(Draught..m.) +
                      log(MRV.Load),
                      # log(Average.Load),
                      log(TPC),
                      log(Service.Speed..knots.),
                      log(Consumption..tpd.))
                  | Size.Category,
               data = AIS_stats_best_df)
etable_iid(models)
etable_cluster(models, 'Built.Cohort')
```

## Interaction
```{r}
AIS_stats_best_df$Built.Cohort = cut(AIS_stats_best_df$Built.Year,
                                     breaks = seq(min(AIS_stats_best_df$Built.Year), 2021, by = 10))

models <- feols(log(FC.Per.Travel.Work) ~ 
                  csw(Age:Built.Cohort + 
                      log(Main.Engine.Power.kW) +
                      log(Dwt) +
                      log(LBP..m.) +
                      log(Beam.Mld..m.) +
                      log(Draught..m.),
                      log(MRV.Load),
                      # log(Average.Load),
                      log(Service.Speed..knots.),
                      log(Consumption..tpd.))
                  | csw(Size.Category, Built.Cohort),
               data = AIS_stats_best_df %>% mutate())
# etable_cluster(models, 'Built.Cohort')
etable_iid(models)
```

```{r}
AIS_stats_best_df$Built.Cohort = cut(AIS_stats_best_df$Built.Year,
                                     breaks = seq(min(AIS_stats_best_df$Built.Year), 2021, by = 9))

models <- feols(log(FC.Per.Travel.Work) ~ 
                  csw(Age:Built.Cohort + 
                      log(Main.Engine.Power.kW) +
                      log(Dwt) +
                      log(LBP..m.) +
                      log(Beam.Mld..m.) +
                      log(Draught..m.) +
                      log(TPC),
                      log(MRV.Load),
                      # log(Average.Load),
                      log(Service.Speed..knots.),
                      log(Consumption..tpd.))
                  | Size.Category^Built.Cohort,
               data = AIS_stats_best_df %>% mutate())
etable_iid(models)
```

# Subset recent ships
2019 reporting period
```{r}
models <- feols(log(FC.Per.Travel.Work) ~ 
  csw(Age +
      log(Main.Engine.Power.kW) +
      log(Dwt) +
      log(LBP..m.) +
      log(Beam.Mld..m.) +
      log(Draught..m.) +
      log(MRV.Load),
      # log(Average.Load),
      log(TPC),
      log(Service.Speed..knots.),
      log(Consumption..tpd.))
  | sw0(Size.Category),
               data = AIS_stats_best_df %>% filter(Year == 2019))
etable_cluster(models, 'Size.Category')
```

2020 reporting period
```{r}
models <- feols(log(FC.Per.Travel.Work) ~ 
  csw(Age +
      log(Main.Engine.Power.kW) +
      log(Dwt) +
      log(LBP..m.) +
      log(Beam.Mld..m.) +
      log(Draught..m.) +
      log(MRV.Load),
      # log(Average.Load),
      log(TPC),
      log(Service.Speed..knots.),
      log(Consumption..tpd.))
  | sw0(Size.Category),
               data = AIS_stats_best_df %>% filter(Year == 2020))
etable_cluster(models, 'Size.Category')
```

Recent ships
```{r}
models <- feols(log(FC.Per.Travel.Work) ~ 
  csw(Age +
      log(Main.Engine.Power.kW) +
      log(Dwt) +
      log(LBP..m.) +
      log(Beam.Mld..m.) +
      log(Draught..m.) +
      log(MRV.Load),
      log(TPC),
      # log(Average.Load),
      log(Service.Speed..knots.),
      log(Consumption..tpd.))
  | sw0(Size.Category),
               data = AIS_stats_best_df %>% filter(Built.Year >= 1996))
etable_cluster(models, 'Size.Category')
```


```{r}
models <- feols(log(FC.Per.Travel.Work) ~ 
  csw(Age +
      log(Main.Engine.Power.kW) +
      log(Dwt) +
      log(LBP..m.) +
      log(Beam.Mld..m.) +
      log(Draught..m.) +
      log(MRV.Load),
      log(TPC),
      # log(Average.Load),
      log(Service.Speed..knots.),
      log(Consumption..tpd.))
  | sw0(Size.Category),
               data = AIS_stats_best_df %>% filter(Built.Year >= 2001))
etable_cluster(models, 'Size.Category')
```

```{r}
models <- feols(log(FC.Per.Travel.Work) ~ 
  csw(Age +
      log(Main.Engine.Power.kW) +
      log(Dwt) +
      log(LBP..m.) +
      log(Beam.Mld..m.) +
      log(Draught..m.) +
      log(MRV.Load),
      log(TPC),
      # log(Average.Load),
      log(Service.Speed..knots.),
      log(Consumption..tpd.))
  | sw0(Size.Category),
               data = AIS_stats_best_df %>% filter(Built.Year >= 2006))
etable_cluster(models, 'Size.Category')
```

```{r}
models <- feols(log(FC.Per.Travel.Work) ~ 
  csw(Age +
      log(Main.Engine.Power.kW) +
      log(Dwt) +
      log(LBP..m.) +
      log(Beam.Mld..m.) +
      log(Draught..m.) +
      log(MRV.Load),
      log(TPC),
      # log(Average.Load),
      log(Service.Speed..knots.),
      log(Consumption..tpd.))
  | sw0(Size.Category),
               data = AIS_stats_best_df %>% filter(Built.Year >= 2011))
etable_cluster(models, 'Size.Category')
```
# END
