---
title: "AIS Data IDs"
output: html_document
---

Create csv of corrected IMO numbers from AIS data
Truncate and validate with checksum as per Fourth IMO report
Input(s): ais_ids.csv
Output(s): ais_corrected_imo.csv


# START
```{r}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
```

# Functions
```{r}
IMO_checksum <- function(imo){
  digits_sum = 0
  n_digits = nchar(imo)
  if((n_digits > 7) | (imo == 0) | (imo == 1234567)){
    valid = FALSE
  }else{
    imo = str_pad(imo, width = 7, side = 'left', pad = '0')
    for(i in 1:6){
      digits_sum = digits_sum + (8-i)*as.integer(str_sub(imo, i, i))
    }
    digits_sum = as.character(digits_sum)
    valid = str_sub(imo, 7, 7) == str_sub(digits_sum, -1)
  }
  return(valid)
}

# min.na <- function(x){
#   ifelse(all(is.na(x)), NA_real_, min(x, na.rm = TRUE))
# }
# 
# max.na <- function(x){
#   ifelse(all(is.na(x), NA_real_, max(x, na.rm = TRUE)))
# }
```

# Read file
```{r}
ids_df <- read_csv("data/ais_ids.csv",
                   col_types = "iddci") %>% 
  rename(Obs.Count = name...5,
         name = name...4)

# load("data/All_join.Rda")
```

IMO number seems to have digits after decimal. Check these are all zero
```{r}
if(!all(ids_df$imo %% 1 == 0)){
  stop('IMO numbers are not integers!')
}else{
  ids_df <- ids_df |> 
    mutate(imo = as.integer(imo))
}
```

# Raw data summary stats
How many mmsi?
```{r}
ids_df |> distinct(mmsi) |> nrow()
```

How many mmsi-imo combinations?
```{r}
ids_df |> 
  group_by(mmsi, imo) |> 
  summarise(Obs.Count = sum(Obs.Count))
```

How many mmsi associated with each imo?
```{r}
ids_df |> 
  filter(imo != 0) |> 
  group_by(mmsi, imo) |> 
  summarise(Obs.Count = sum(Obs.Count)) |> 
  filter(Obs.Count > 10) |> 
  group_by(mmsi) |> 
  mutate(n = n()) |> 
  filter(n == 2) |>
  arrange(mmsi)
```

## IMO checksum
Function check
```{r}
if(IMO_checksum(9834203)){ #invalid
  stop('checksum function not working')
}
if(!IMO_checksum(9074729)){ #valid (example from https://en.wikipedia.org/wiki/IMO_number)
  stop('checksum function not working')
}

```

Check all IMO
```{r}
ids_df <- ids_df |> 
  rowwise() |> 
  mutate(imo.isvalid = IMO_checksum(imo)) |> 
  ungroup()
```

How many valid imo's per mmsi?
```{r}
ids_df |> 
  group_by(mmsi) |>
  distinct(imo, imo.isvalid) |> 
  summarise(valid.imos = sum(imo.isvalid, na.rm = TRUE)) |> 
  count(valid.imos)
```
Almost 3000 have two valid IMO numbers, 651 have three.

Which have many? (seems implausible)
```{r}
ids_df |> 
  group_by(mmsi) |>
  distinct(imo, imo.isvalid) |> 
  summarise(valid.imos = sum(imo.isvalid, na.rm = TRUE)) |> 
  filter(valid.imos == 5)
```

Any better if I fix as IMO does, removing extra digits?
```{r}
ids_df <- ids_df |> 
  mutate(imo.truncated = ifelse(nchar(imo) > 7, as.integer(imo/10^(nchar(imo)-7)), imo)) |> 
  rowwise() |>
  mutate(imo.truncated.isvalid = IMO_checksum(imo.truncated)) |> 
  ungroup()
```

How many valid imo's per mmsi after truncation?
```{r}
ids_df |> 
  group_by(mmsi) |>
  distinct(imo.truncated, imo.truncated.isvalid) |> 
  summarise(valid.truncated.imos = sum(imo.truncated.isvalid, na.rm = TRUE)) |> 
  count(valid.truncated.imos)
```

How many fixed?
```{r}
ids_df |> 
  filter(imo.isvalid != imo.truncated.isvalid) |> 
  count(imo.isvalid, imo.truncated.isvalid)
```
2169 rows!

```{r}
ids_df <- ids_df |> 
  mutate(imo.corrected = ifelse(imo.truncated.isvalid, imo.truncated, NA_integer_))
```

# Save corrected list
```{r}
mmsi_imo_df <- ids_df |> 
  distinct(mmsi, imo, imo_corrected = imo.corrected) 

mmsi_imo_df |> 
  write_csv("data/ais_corrected_imo.csv")
```

