---
title: "AIS Data IDs"
output: html_document
---

# START
```{r}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library("tabulapdf")
```

# Functions
```{r}
IMO_checksum <- function(imo){
  digits_sum = 0
  n_digits = nchar(imo)
  if((n_digits > 7) | (imo == 0) | (imo == 1234567)){
    valid = FALSE
  }else{
    imo = str_pad(imo, width = 7, side = 'left', pad = '0')
    for(i in 1:6){
      digits_sum = digits_sum + (8-i)*as.integer(str_sub(imo, i, i))
    }
    digits_sum = as.character(digits_sum)
    valid = str_sub(imo, 7, 7) == str_sub(digits_sum, -1)
  }
  return(valid)
}

# min.na <- function(x){
#   ifelse(all(is.na(x)), NA_real_, min(x, na.rm = TRUE))
# }
# 
# max.na <- function(x){
#   ifelse(all(is.na(x), NA_real_, max(x, na.rm = TRUE)))
# }
```

# Read file
```{r}
ids_df <- read_csv("data/ais_ids.csv",
                   col_types = "iddci") %>% 
  rename(Obs.Count = name...5,
         name = name...4)

# load("data/All_join.Rda")
```

IMO number seems to have digits after decimal. Check these are all zero
```{r}
if(!all(ids_df$imo %% 1 == 0)){
  stop('IMO numbers are not integers!')
}else{
  ids_df <- ids_df |> 
    mutate(imo = as.integer(imo))
}
```

# Raw data summary stats
How many mmsi?
```{r}
ids_df |> distinct(mmsi) |> nrow()
```

How many mmsi-imo combinations?
```{r}
ids_df |> 
  group_by(mmsi, imo) |> 
  summarise(Obs.Count = sum(Obs.Count))
```

How many mmsi associated with each imo?
```{r}
ids_df |> 
  filter(imo != 0) |> 
  group_by(mmsi, imo) |> 
  summarise(Obs.Count = sum(Obs.Count)) |> 
  filter(Obs.Count > 10) |> 
  group_by(mmsi) |> 
  mutate(n = n()) |> 
  filter(n == 2) |>
  arrange(mmsi)
```

## IMO checksum
Function check
```{r}
if(IMO_checksum(9834203)){ #invalid
  stop('checksum function not working')
}
if(!IMO_checksum(9074729)){ #valid (example from https://en.wikipedia.org/wiki/IMO_number)
  stop('checksum function not working')
}

```

Check all IMO
```{r}
ids_df <- ids_df |> 
  rowwise() |> 
  mutate(imo.isvalid = IMO_checksum(imo)) |> 
  ungroup()
```

How many valid imo's per mmsi?
```{r}
ids_df |> 
  group_by(mmsi) |>
  distinct(imo, imo.isvalid) |> 
  summarise(valid.imos = sum(imo.isvalid, na.rm = TRUE)) |> 
  count(valid.imos)
```
Almost 3000 have two valid IMO numbers, 651 have three.

Which have many? (seems implausible)
```{r}
ids_df |> 
  group_by(mmsi) |>
  distinct(imo, imo.isvalid) |> 
  summarise(valid.imos = sum(imo.isvalid, na.rm = TRUE)) |> 
  filter(valid.imos == 5)
```

Any better if I fix as IMO does, removing extra digits?
```{r}
ids_df <- ids_df |> 
  mutate(imo.truncated = ifelse(nchar(imo) > 7, as.integer(imo/10^(nchar(imo)-7)), imo)) |> 
  rowwise() |>
  mutate(imo.truncated.isvalid = IMO_checksum(imo.truncated)) |> 
  ungroup()
```

How many valid imo's per mmsi after truncation?
```{r}
ids_df |> 
  group_by(mmsi) |>
  distinct(imo.truncated, imo.truncated.isvalid) |> 
  summarise(valid.truncated.imos = sum(imo.truncated.isvalid, na.rm = TRUE)) |> 
  count(valid.truncated.imos)
```

How many fixed?
```{r}
ids_df |> 
  filter(imo.isvalid != imo.truncated.isvalid) |> 
  count(imo.isvalid, imo.truncated.isvalid)
```
2169 rows!

```{r}
ids_df <- ids_df |> 
  mutate(imo.corrected = ifelse(imo.truncated.isvalid, imo.truncated, NA_integer_))
```

# Save corrected list
```{r}
mmsi_imo_df <- ids_df |> 
  distinct(mmsi, imo, imo_corrected = imo.corrected) 

mmsi_imo_df |> 
  write_csv("data/ais_corrected_imo.csv")
```


# Load WFR
## Functions

```{r}
extract_fleet_page <- function(x){
  df <- str_split(x, '\n')[[1]] %>% 
    as_tibble()
  
  df_out <- df %>%
    rowid_to_column('Row') %>%
    filter(str_detect(value, 'DWT|built'))
  # second_rows <- df[df_out$Row + 1,]
  # colnames(second_rows) <- 'Second.Row'
  imo_rows <- df %>% 
    filter(str_detect(value, 'IMO')) %>% 
    rename(IMO.Row = value)
  mmsi_rows <- df |> 
    filter(str_detect(value, 'MMSI')) |> 
    rename(MMSI.Row = value)
  df_out <- bind_cols(df_out, imo_rows, mmsi_rows) %>% 
    select(-Row)
  return(df_out)
}
```

```{r}
extract_fleet_register <- function(scrape){
  register_df <- lapply(scrape, extract_fleet_page) %>% 
    bind_rows(.id = "Page") %>% 
    mutate(value = ifelse(value == "Bulk Carrier, 74,483 DWT, built 5/2006 at Hudong Zhonghua , IMO 9358838 ",
                      "Nan Xin 27  Bulk Carrier, 74,483 DWT, built 5/2006 at Hudong Zhonghua , IMO 9358838",
                      value),
       value = ifelse(value == "Open Hatch Carrier, 32,259 DWT, built 2/2002 at Saiki Hvy. Ind. , IMO 9246308 ",
                      "Singapore Spirit  Open Hatch Carrier, 32,259 DWT, built 2/2002 at Saiki Hvy. Ind. , IMO 9246308",
                      value),
       value = ifelse(value == "Bulk Carrier, 81,702 DWT, built 11/2008 at Mitsui SB (Tamano) , IMO 9389239 ",
                      "YM Effort  Bulk Carrier, 81,702 DWT, built 11/2008 at Mitsui SB (Tamano) , IMO 9389239",
                      value)) %>% 
    mutate(IMO = str_extract(IMO.Row, '(?<=IMO\\s)[:digit:]+'),
           MMSI = str_extract(MMSI.Row, '(?<=MMSI\\s)[:digit:]+')) %>% 
    select(-IMO.Row, -MMSI.Row) %>% 
  return(register_df)
}
```

```{r}
extract_fleet_registers <- function (row, page_index){
  year <- page_index$Issue[row]
  start_page <- page_index$Start.Page[row]
  end_page <- page_index$End.Page[row]
  file_prefix <- page_index$File.Prefix[row]
  print(year)
  register_df <- extract_text(file.path(registerpath, paste0(file_prefix, year, '.pdf')),
                                pages = start_page:end_page) %>% 
    # str_remove_all(paste0('(\nOwned/Managed by|\nLast ex name( |):) .* \n(?=', shiptypes, ')')) %>% 
    extract_fleet_register() %>% 
    mutate(Issue = year)
  return(register_df)
}
```


## Extract IDs from BCR pdfs
```{r}
registerpath <- file.path("data", "WFR", "Bulk_Carrier_Register")
```

```{r}
page_index <- tribble(
  ~Issue, ~Start.Page, ~End.Page, ~Split, ~OCR, ~File.Prefix,
    # 2006,         171,       873, FALSE, FALSE, 'The Bulk Carrier Register ',
    # 2013,         204,      1271, FALSE, FALSE, 'The Bulk Carrier Register ',
    # 2014,         202,      1316, FALSE, FALSE, 'The Bulk Carrier Register ',
    # 2015,         200,      1358, TRUE,  TRUE, 'BCR_fleet_OCR_',
    # 2016,         196,      1391, FALSE, FALSE, 'The Bulk Carrier Register ',
    # 2017,         195,      1419, FALSE, FALSE, 'The Bulk Carrier Register ',
    # 2018,         191,      1216, FALSE, FALSE, 'The Bulk Carrier Register ',
    # 2019,         195,      1227, TRUE,  TRUE, 'BCR_fleet_OCR_',
    # 2019,         195,      1227, FALSE,  FALSE, 'The Bulk Carrier Register '
    2019,             1,      1033, FALSE,  FALSE, 'BCR_fleet_',
    2020,         207,      1204, FALSE, FALSE, 'The Bulk Carrier Register ',
    2021,         205,      1231, FALSE, FALSE, 'The Bulk Carrier Register ',
    2022,         225,      1283, FALSE, FALSE, 'The Bulk Carrier Register ',
)

page_index <- page_index |> 
  rowwise() |> 
  mutate(n.files = list.files(registerpath) |> 
           str_detect(paste0(File.Prefix, '[:digit:]_')) |> 
           which() |>
           length()) |> 
  ungroup() |> 
  mutate(n.files = replace(n.files, n.files == 0, 1)) |> 
  uncount(n.files) |> 
  group_by(Issue) |> 
  mutate(split.index = row_number(),
         File.Prefix = replace(File.Prefix, Split, paste0(File.Prefix, split.index, '_')),
         Start.Page = replace(Start.Page, Split, 1),
         End.Page = replace(End.Page, Split, 250)) |> 
  select(-split.index)

page_index
```

```{r}
register_df <- lapply(1:nrow(page_index),
                      extract_fleet_registers, page_index = page_index) %>% 
  bind_rows()
```

```{r}
save(register_df, file = "data/wfr_bulkers_mmsi_register.Rda")
load("data/wfr_bulkers_mmsi_register.Rda")
```


How many ships per year?
```{r}
register_df |> 
  select(Issue, Page, IMO, MMSI) |>
  count(Issue)
```

How many have multiple mmsi?
```{r}
register_df |> 
  select(Issue, Page, IMO, MMSI) |>
  drop_na(IMO, MMSI) |> 
  distinct(IMO, MMSI) |> 
  count(IMO, name = 'n.MMSI') |>
  count(n.MMSI)
```
About 11%

```{r}
register_df |> 
  select(Issue, Page, IMO, MMSI) |>
  drop_na(IMO, MMSI) |> 
  distinct(IMO, MMSI) |>
  group_by(IMO) |> 
  mutate(n.MMSI = n()) |> 
  arrange(IMO)
```

### MMSI-based
How many imo per mmsi?
```{r}
register_df |> 
  select(Issue, Page, IMO, MMSI) |>
  drop_na(IMO, MMSI) |> 
  distinct(IMO, MMSI) |> 
  count(MMSI, name = 'n.IMO') |>
  count(n.IMO)
```
There shouldn't be many with multiple IMO per MMSI

```{r}
mmsi_imo_wfr_df <- register_df |> 
  select(Issue, Page, IMO, MMSI) |>
  drop_na(IMO, MMSI) |> 
  distinct(IMO, MMSI) |> 
  mutate(across(everything(), as.integer))
```


# Join and Compare
```{r}
mmsi_imo_ais_df <- read_csv("data/ais_corrected_imo.csv")
mmsi_imo_ais_df <- mmsi_imo_ais_df |> 
  distinct(mmsi, imo_corrected) |> 
  drop_na(imo_corrected)
mmsi_imo_ais_df
```
How many valid imo per mmsi?
```{r}
mmsi_imo_ais_df |> 
  count(mmsi, name = 'n.IMO') |> 
  count(n.IMO)
```

How many valid mmsi per imo?
```{r}
mmsi_imo_ais_df |> 
  count(imo_corrected, name = 'n.MMSI') |> 
  count(n.MMSI)
```

```{r}
mmsi_imo_compare_df <- full_join(mmsi_imo_ais_df |> rename(imo = imo_corrected) |> mutate(ais = TRUE),
          mmsi_imo_wfr_df |> rename(imo = IMO, mmsi = MMSI) |>  mutate(wfr = TRUE),
          by = c('mmsi', 'imo')) |> 
  mutate(across(c(ais, wfr), \(x) replace_na(x, FALSE))) |> 
  mutate(both = ais & wfr)
```

```{r}
mmsi_imo_compare_df |> 
  count(ais, wfr)
```
```{r}
mmsi_imo_compare_df |> 
  arrange(mmsi) |> 
  group_by(mmsi) |> 
  summarise(n.ais = sum(ais, na.rm = TRUE),
            n.wfr = sum(wfr, na.rm = TRUE),
            n.both = sum(both, na.rm = TRUE),
            n = n(),
            all.match = n.both == n,
            .groups = 'drop') |> 
  count(all.match)
```


What would I get the original way (on WFR mmsi from 2021 only)? (not sure if I used 2021)
```{r}
imo_mmsi_wfr_2021_df <- register_df |> 
  filter(Issue == 2022) |> 
  select(IMO, MMSI)
```

```{r}
mmsi_imo_ais_df |> 
  filter(mmsi %in% imo_mmsi_wfr_2021_df$MMSI) |> 
  count(mmsi, name = "n.imo") |> 
  count(n.imo)
```
With 2021, should cleanly get 7424 (1 IMO associated), and some fraction of the 2722 that have 2 IMO's associated.
With 2022, should cleanly get 7919, and some of 2567 with 2 IMO's.

How many could I potentially improve matching for?

Strategy 1:
1. match on mmsi and year
```{r}
mmsi_imo_ais_df |> 
  filter(mmsi %in% mmsi_imo_wfr_df$MMSI) |> 
  count(mmsi, name = "n.imo") |> 
  count(n.imo)
```
Almost 1000 more clean matches! This could be worthwhile.
```{r}
(1 - 7424/8366)*100
```

However, need to detect when the change occurs.

To get a sense of how many mmsi with multiple imo's associated in tracking data that I might be able to separate,
can try to check contiguous imo numbers. Do this in AIS_Data_IDs.py

# Check ordered
```{r}
mmsi_imo_ais_ordered_df <- read_csv("data/contig_obs.csv")
mmsi_imo_ais_ordered_df 
mmsi_imo_ais_ordered_df |> filter(imo_corrected == 9168348)
```

```{r}
mmsi_imo_ais_ordered_df |> 
  relocate(imo_corrected, .before = mmsi) |> 
  group_by(mmsi) |> 
  arrange(imo_corrected, first_obs) |> 
  View()
```


```{r}
imo_mmsi_ais_df <- mmsi_imo_ais_ordered_df |> 
  group_by(mmsi) |> 
  filter(all(n_instances == 1)) |>
  arrange(imo_corrected, first_obs) |>
  select(imo_corrected, mmsi) |> 
  group_by(imo_corrected) |> 
  summarise(mmsi.list = list(mmsi))
```


### Join with WFR
```{r}
imo_mmsi_wfr_df <- register_df |> 
  select(Issue, IMO, MMSI) |>
  drop_na(IMO, MMSI) |> 
  arrange(Issue) |> 
  distinct(IMO, MMSI) |> 
  mutate(across(everything(), as.integer)) |> 
  group_by(IMO) |>
  summarise(mmsi.list = list(MMSI))
imo_mmsi_wfr_df
```


```{r}
# imo_mmsi_wfr_df <- 
mmsi_imo_wfr_df #|>
  # group_by(IMO) |>
  # summarise(mmsi.list = list(MMSI))
```

```{r}
imo_mmsi_ordered_compare_df <- full_join(imo_mmsi_ais_df,
                                         imo_mmsi_wfr_df,
                                         by = c('imo_corrected' = 'IMO'),
                                         suffix = c('.ais', '.wfr')) |> 
  rowwise() |> 
  mutate(ais.notin.wfr = sum(!(mmsi.list.ais %in% mmsi.list.wfr)),
         wfr.in.ais = sum(mmsi.list.wfr %in% mmsi.list.ais),
         ais.in.wfr = sum(mmsi.list.ais %in% mmsi.list.wfr),
         same.length = length(mmsi.list.ais) == length(mmsi.list.wfr),
         all.ais.in.wfr = all(mmsi.list.ais %in% mmsi.list.wfr)) |> 
  ungroup()

imo_mmsi_ordered_compare_df
```

Probably hopeless if many ais mmsi not in wfr
```{r}
imo_mmsi_ordered_compare_df |> 
  count(ais.notin.wfr)
```
Most are accounted for actually!


```{r}
imo_mmsi_ordered_compare_df |> 
  count(wfr.in.ais, ais.in.wfr, same.length, all.ais.in.wfr) |> 
  arrange(all.ais.in.wfr, same.length)
```

```{r}
imo_mmsi_ordered_compare_df |> 
  count(all.ais.in.wfr) |> 
  arrange(all.ais.in.wfr)
```


It seems there is progress to be made here.
But how to split AIS paths at the proper time?

Strategy 1: 
  -x find start and end dates in static data
  - look for nearest begin, end, or jump in mmsi paths and split here (create path column)
    - how best to look for these points?
    - start by using current path splitting function to create df of jump times
  - assign IMO (most frequent?) and re-index to it
  - re-sort by IMO and timestamp
  - remove skips and jumps as before
