---
title: "All Join"
output: html_document
---

```{r}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
# library(feather)
```

# Load data
```{r}
callvariant = 'speed' #'heading'
EUvariant = '_EEZ' #''


load("data/bulkers_WFR.Rda")
load("data/containerships_WFR.Rda")
load("data/oiltankers_WFR.Rda")

AIS_df <- read.csv("data/AIS_yearly_stats.csv") %>%
  as_tibble(.name_repair = 'universal')
colnames(AIS_df) <- colnames(AIS_df) %>%
  str_replace_all('_', ' ') %>%
  str_to_title() %>%
  str_replace_all(' ', '.')

AIS_EU_df <- read_csv(paste0("data/AIS_", callvariant, EUvariant, "_EU_yearly_stats.csv")) %>% 
  as_tibble(.name_repair = 'universal') %>%
  rename(MMSI = mmsi,
         Distance = distance_sum,
         Work = work_sum,
         Work.IS = work_IS_sum,
         Trips = trip_nunique
         # InvalidSpeed = invalid_speed
         )

load("data/MRV.Rda")
```

Harmonize types
```{r}
bulkers_df <- bulkers_df %>% 
  mutate(IMO.Number = as.integer(IMO.Number))

containerships_df <- containerships_df %>% 
  mutate(IMO.Number = as.integer(IMO.Number))

oiltankers_df <- oiltankers_df %>% 
  mutate(IMO.Number = as.integer(IMO.Number))

MRV_df <- MRV_df %>% 
  mutate(imo.number = as.integer(imo.number))
```

# EIV Calculations
## Bulkers
```{r}
bulkers_df <- bulkers_df %>% 
  mutate(Main.Engine.Power.kW = str_remove_all(Main.Engine.Detail, ",") %>% 
                             str_extract("\\b\\d*(?=mkW\\stotal)") %>% 
                             as.integer(),
         Auxiliary.Engine.Power.kW = str_remove_all(Auxiliary.Engine.Detail, ",") %>% 
                             str_extract("\\b\\d*(?=mkW\\stotal)") %>% 
                             as.integer()) %>% 
  mutate(PAE = if_else(Main.Engine.Power.kW >= 10000, 0.025 * Main.Engine.Power.kW + 250, 0.05 * Main.Engine.Power.kW),
         Design.Efficiency = Main.Engine.Power.kW*Main.Consumption.at.Service.Speed..tpd../(Dwt*Service.Speed..knots.),
         EIV.Calculated = 3.1144 * (190 * 0.75 * Main.Engine.Power.kW + 215 * PAE)/(Dwt*Service.Speed..knots.),
         EIV.Main.Calculated = 3.1144 * (190 * 0.75 * Main.Engine.Power.kW)/(Dwt*Service.Speed..knots.))
```

```{r}
bulkers_df %>% 
  ggplot(aes(PAE)) + 
  geom_histogram(binwidth = 20)
```

```{r}
bulkers_df %>% 
  ggplot(aes(x = EIV.Calculated, fill = Size.Category)) + 
  geom_histogram(binwidth = 0.25)
```

# Combine ship types
```{r}
fleet_df <- bind_rows(bulkers_df, containerships_df, oiltankers_df)
```

## Reorganize
```{r}
fleet_df <- fleet_df %>% 
  relocate(Fleet.Type, .before = CVN) %>% 
  relocate(c(Size.Category, Dwt), .after = MMSI) %>% 
  relocate(Type, .before = Size.Category) %>% 
  # relocate(Dwt, .after = Size.Category) %>% 
  relocate(c(Built.Date, Built.Year), .after = Dwt) %>% 
  relocate(Alternative.Name, .after = Name) %>% 
  relocate(Ex.Name, .after = Alternative.Name) %>% 
  relocate(Owner.Size, .after = Owner) %>% 
  relocate(Design.Efficiency, .after = Group.Owner) %>% 
  relocate(contains("EIV"), .after = Design.Efficiency)
```


# WFR-AIS
```{r}
AIS_df <- AIS_df %>% 
  rename(MMSI = Mmsi) %>% 
  arrange(Year) %>% 
  pivot_wider(names_from = Year,
              names_sep = ".",
              values_from = -c('MMSI', 'Year'))
```

```{r}
AIS_EU_df <- AIS_EU_df %>%
  arrange(year) %>% 
  pivot_longer(cols = -c(MMSI, year)) %>% 
  pivot_wider(names_from = c(name, year),
              names_prefix = 'EU.',
              names_sep = '.',
              values_from = value)
```

```{r}
AIS_df <- AIS_df %>% 
  left_join(AIS_EU_df, by = "MMSI")

colnames(AIS_df) <- str_replace(colnames(AIS_df), 'Distance.Sum', 'Distance.Total')
colnames(AIS_df) <- paste0('AIS.', colnames(AIS_df))
```

## join
```{r}
fleet_df <- fleet_df %>% 
  left_join(AIS_df,
            by = c("MMSI" = "AIS.MMSI"),
            keep = TRUE) %>% 
  relocate(AIS.MMSI, .after = MMSI) %>% 
  relocate(contains("Distance"), .after = Group.Owner)
```

How many ships in WFR don't have tracking data?
```{r}
fleet_df %>% 
  drop_na(MMSI) %>% 
  group_by(Fleet.Type, is.na(AIS.MMSI)) %>% 
  summarise(n())
```
How many AIS.MMSI aren't unique?
```{r}
fleet_df %>% 
  drop_na(AIS.MMSI) %>% 
  group_by(AIS.MMSI) %>% 
  mutate(count = n()) %>% 
  filter(count > 1) %>% 
  arrange(AIS.MMSI)
```

Which ships don't have tracking data?
```{r}
fleet_df %>% 
  drop_na(MMSI) %>% 
  group_by(Fleet.Type, is.na(AIS.MMSI), Status) %>% 
  summarise(n())
```
Only missing 185 in-service bulkers! There is some mislabeling or maybe transferred tracking devices, but it's in the very low tens.

Which tracked ships aren't in fleet register?
```{r}
unmatched_ais_df <- anti_join(AIS_df, fleet_df,
                              by = c("AIS.MMSI" = "MMSI"))
```
This will be empty because I used the fleet register to subset AIS data.

```{r eval = FALSE}
AIS_df %>% 
  arrange(desc(AIS.MMSI)) %>% 
  View()
```


# WFR-MRV
```{r eval = FALSE}
MRV_df %>% colnames()
```


```{r}
MRV_wide_df <- MRV_df %>% 
  select(IMO.Number = imo.number,
         Ship.Name = ship.name,
         Ship.Type = ship.type,
         Technical.Efficiency = technical.efficiency,
         Efficiency.Metric = efficiency.metric,
         EEDI,
         EIV,
         Reporting.Period = reporting.period,
         Total.FC = total.fc,
         Total.CO2 = total.co2,
         Time.At.Sea = time.at.sea,
         Speed = speed,
         EU.Distance = EU.distance,
         Emission.Factor = emission.factor,
         Load = load,
         FC.On.Laden = fc.on.laden,
         FC.Per.Distance = fc.per.distance,
         FC.Per.Work.Mass = fc.per.work.mass,
         FC.Per.Work.Mass.On.Laden.Voyages = fc.per.work.mass.on.laden.voyages,
         FC.Per.Work.Volume.On.Laden.Voyages = fc.per.work.volume.on.laden.voyages) %>% 
  arrange(IMO.Number, Reporting.Period) %>%
  pivot_wider(id_cols = c(IMO.Number),
              names_from = Reporting.Period,
              values_from = -c(IMO.Number, Reporting.Period),
              names_sep = ".")

colnames(MRV_wide_df) <- paste0('MRV.', colnames(MRV_wide_df))
```

Can I reduce design efficiency metrics to single variable?
How many ships report the same efficiency year on year?
```{r}
MRV_wide_df %>% 
  mutate(MRV.Constant.Efficiency.Metric = 
           MRV.Efficiency.Metric.2018 == MRV.Efficiency.Metric.2019 &
           MRV.Efficiency.Metric.2019 == MRV.Efficiency.Metric.2020 &
           MRV.Efficiency.Metric.2020 == MRV.Efficiency.Metric.2021,
         MRV.Constant.Efficiency = 
           MRV.Technical.Efficiency.2018 - MRV.Technical.Efficiency.2019 < 1 &
           MRV.Technical.Efficiency.2019 - MRV.Technical.Efficiency.2020 < 1 &
           MRV.Technical.Efficiency.2020 - MRV.Technical.Efficiency.2021 < 1) %>% 
  group_by(MRV.Constant.Efficiency.Metric, MRV.Constant.Efficiency) %>% 
  # filter(constant.efficiency.metric, !constant.efficiency) %>% 
  summarise(n())
```
Leave as-is for now.

Are all MRV reporting bulkers in WFR?
```{r}
anti_join(MRV_wide_df, fleet_df, by = c("MRV.IMO.Number" = "IMO.Number")) %>% 
  group_by(MRV.Ship.Type.2018, MRV.Ship.Type.2019, MRV.Ship.Type.2020, MRV.Ship.Type.2021) %>% 
    filter(MRV.Ship.Type.2018 == 'Bulk carrier' | 
         MRV.Ship.Type.2018 == 'Bulk carrier' | 
         MRV.Ship.Type.2018 == 'Bulk carrier' |
         (is.na(MRV.Ship.Type.2018) & is.na(MRV.Ship.Type.2019) & is.na(MRV.Ship.Type.2020) & is.na(MRV.Ship.Type.2021))) %>% 
  summarise(n())
```
Not many bulkers getting lost!    


## join
```{r}
fleet_df <- fleet_df %>% 
  left_join(MRV_wide_df,
            by = c("IMO.Number" = "MRV.IMO.Number"),
            keep = TRUE) %>% 
  relocate(MRV.IMO.Number, .after = IMO.Number) %>% 
  relocate(contains("MRV.Ship.Name."), .after = Name) %>% 
  relocate(contains("MRV.Ship.Type."), .after = Type) %>% 
  relocate(contains("EU.Distance"), .after = AIS.Distance.Total.2021) %>%
  relocate(contains("EIV"), .after = Design.Efficiency) %>% 
  relocate(contains("EEDI"), .after = MRV.EIV.2021)
```

How many WFR ships report in MRV?
```{r}
fleet_df %>% 
  select(CVN, 
         IMO.Number, 
         MRV.IMO.Number,
         Name, 
         starts_with("MRV.Ship.Name."),
         Type, 
         starts_with("MRV.Ship.Type.")) %>% 
  drop_na(IMO.Number) %>%
  group_by(is.na(MRV.IMO.Number)) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = count / sum(count) * 100)
```
41% of ships report in MRV. This is quite high!

```{r eval = FALSE}
fleet_df %>% 
  select(CVN, 
         IMO.Number, 
         MRV.IMO.Number,
         Name, 
         starts_with("MRV.Ship.Name."),
         Type, 
         starts_with("MRV.Ship.Type.")) %>% 
  filter(!is.na(MRV.IMO.Number)) %>% 
  View()
```

How does fleet type compare between WFR and MRV?
```{r}
fleet_df %>% 
  group_by(Fleet.Type, MRV.Ship.Type.2020) %>% 
  summarise(n())
```
Fairly good correspondence, though MRV has many missings.


# WFR-AIS-MRV
```{r eval = FALSE}
fleet_df %>% 
  group_by(Fleet.Type, is.na(AIS.MMSI), is.na(MRV.IMO.Number)) %>% 
  summarise(n())
```


# Save
```{r}
save(fleet_df, file = "data/All_join.Rda")
```

