---
title: "MRV-WFR Exploration"
output: html_document
---
```{r}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(stargazer)
```


```{r}
load("data/All_join.Rda")

fleet_df$Owner.Size <- factor(fleet_df$Owner.Size,
                                c("Very Small (1-5)",
                                    "Small (6-10)",
                                    "Medium (11-20)",
                                    "Large (21-50)",
                                    "Very Large (51-100)",
                                    "Extra Large (100+)"))

```

## MRV coverage
How many report in MRV?
```{r}
fleet_df %>% 
  drop_na(IMO.Number) %>%
  group_by(is.na(MRV.IMO.Number)) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = count / sum(count) * 100)
```

```{r}
fleet_df %>% 
  drop_na(IMO.Number) %>%
  group_by(is.na(MRV.Total.FC.2019)) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = count / sum(count) * 100)
```

By year...
```{r}
fleet_df %>% 
  select(CVN, IMO.Number, MRV.IMO.Number, contains("Total.FC")) %>% 
  drop_na(IMO.Number) %>%
  mutate(across(contains("Total.FC."), ~ !is.na(.x), .names = "Reporting.{.col}")) %>% 
  pivot_longer(starts_with("Reporting.MRV.Total.FC."),
               values_to = "Reporting",
               names_to = "Reporting.Year",
               names_prefix = "Reporting.MRV.Total.FC.") %>% 
  group_by(Reporting.Year, Reporting) %>% 
  summarise(count = n()) %>% 
  mutate(Reporting.Percentage = count/sum(count) * 100)
```
About 30% of all bulker ships report in the MRV each year.

## Check fuel consumption per distance makes sense
```{r}
fleet_df |> 
  select(IMO.Number, contains('MRV.EU.Distance'), contains('MRV.FC.Per.Distance'), contains('MRV.Total.FC')) |>
  select(IMO.Number, contains('2018')) |> 
  mutate(test = MRV.Total.FC.2018/MRV.EU.Distance.2018*1e3 - MRV.FC.Per.Distance.2018) |> 
  drop_na('test') |> 
  summarise(all(test < 1e-2))
```



## Fuel consumption and efficiency
```{r}
fleet_df %>% 
  select(CVN, contains("total.fc"), Built.Date, Size.Category) %>% 
  pivot_longer(contains("total.fc"),
               values_to = "total.fc",
               names_to = "reporting.period") %>% 
  ggplot(aes(y = total.fc, 
             x = Built.Date,
             color = reporting.period)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~Size.Category) +
  coord_cartesian(ylim = c(0, 13000))
```

```{r}
fleet_df %>% 
  select(CVN, contains("fc.per.distance"), Built.Date, Size.Category) %>% 
  pivot_longer(contains("fc.per.distance"),
               values_to = "fc.per.distance",
               names_to = "reporting.period") %>% 
  ggplot(aes(fc.per.distance)) +
  geom_histogram(binwidth = 10) +
  xlim(0, 1000)
```

```{r}
fleet_df %>% 
  select(CVN, contains("fc.per.work.mass"), Built.Date, Size.Category) %>% 
  pivot_longer(contains("fc.per.work.mass"),
               values_to = "fc.per.work.mass",
               names_to = "reporting.period") %>% 
  ggplot(aes(fc.per.work.mass)) +
  geom_histogram(binwidth = 0.5) +
  xlim(0, 30)
```

### Distance vs Work measure
```{r}
fleet_df %>% 
  select(-contains("MRV.fc.per.work.mass.on")) %>% 
  select(CVN, Built.Date, Size.Category, contains("fc.per.distance"), contains("fc.per.work.mass")) %>% 
  pivot_longer(c("MRV.FC.Per.Distance.2018", "MRV.FC.Per.Distance.2019", "MRV.FC.Per.Distance.2020",
                 "MRV.FC.Per.Work.Mass.2018", "MRV.FC.Per.Work.Mass.2019", "MRV.FC.Per.Work.Mass.2020")) %>% 
  separate(name, c("name", "reporting.period"), sep = ".20", ) %>% 
  mutate(reporting.period = 2000 + as.integer(reporting.period)) %>% 
  pivot_wider(names_from = name,
              values_from = value) %>% 
  # group_by(is.na(fc.per.distance), is.na(fc.per.work.mass)) %>% 
  # summarise(n())
  ggplot(aes(y = MRV.FC.Per.Work.Mass, 
             x = MRV.FC.Per.Distance,
             color = Size.Category)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylim(0, 30) +
  xlim(0, 250)
```

```{r}
regression_df <- fleet_df %>% 
  select(-contains("fc.per.work.mass.on")) %>% 
  select(CVN, Dwt, contains("fc.per.distance"), contains("fc.per.work.mass")) %>% 
  pivot_longer(c("MRV.FC.Per.Distance.2018", "MRV.FC.Per.Distance.2019", "MRV.FC.Per.Distance.2020",
                 "MRV.FC.Per.Work.Mass.2018", "MRV.FC.Per.Work.Mass.2019", "MRV.FC.Per.Work.Mass.2020")) %>% 
  separate(name, c("name", "reporting.period"), sep = ".20", ) %>% 
  mutate(reporting.period = 2000 + as.integer(reporting.period)) %>% 
  pivot_wider(names_from = name,
              values_from = value) %>% 
  drop_na(MRV.FC.Per.Distance, MRV.FC.Per.Work.Mass)

lm_fc_1 <- lm(MRV.FC.Per.Work.Mass ~ MRV.FC.Per.Distance + Dwt, regression_df)
lm_fc_2 <- lm(MRV.FC.Per.Work.Mass ~ MRV.FC.Per.Distance + Dwt + I(Dwt^2), regression_df)
lm_fc_3 <- lm(MRV.FC.Per.Work.Mass ~ MRV.FC.Per.Distance*Dwt, regression_df)
stargazer(lm_fc_1, lm_fc_2, lm_fc_3, type = "text")
```



### vs. Age
```{r}
fleet_df %>% 
  select(CVN, contains("fc.per.distance"), Built.Date, Size.Category) %>% 
  pivot_longer(contains("fc.per.distance"),
               values_to = "fc.per.distance",
               names_to = "reporting.period") %>% 
  ggplot(aes(y = fc.per.distance, 
             x = Built.Date,
             color = reporting.period)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~Size.Category) +
  coord_cartesian(ylim = c(0, 1000))
```



```{r}
fleet_df %>% 
  select(-contains("fc.per.work.mass.on")) %>% 
  pivot_longer(contains("fc.per.work.mass"),
               values_to = "fc.per.work.mass",
               names_to = "reporting.period") %>% 
  filter(fc.per.work.mass < 750) %>% # filter outliers
  ggplot(aes(y = fc.per.work.mass, 
             x = Built.Date,
             color = reporting.period)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  # coord_cartesian(ylim = c(0, 100)) +
  ylim(0,20) +
  facet_wrap(~Size.Category)
```

```{r}
fleet_df %>% 
  filter(Built.Year >= 1980) %>% 
  ggplot(aes(y = EIV.Calculated, 
             x = Built.Date)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  # coord_cartesian(ylim = c(0, 100)) +
  ylim(0, 40) +
  facet_wrap(~Size.Category)
```

##### NAMES FIXED TO HERE

### Regressions
```{r}
regression_df <- fleet_df %>% 
  group_by(Owner) %>% 
  mutate(Owner.Size.fleet = n()) %>% 
  ungroup() %>% 
  filter(MRV.FC.Per.Work.Mass.2020 <= 20)

lm_fc_1 <- lm(fc.per.work.mass.2020 ~ design.efficiency,
              data = regression_df %>% 
                select(fc.per.work.mass.2020,
                       design.efficiency) %>% 
                drop_na())
lm_fc_2 <- lm(fc.per.work.mass.2020 ~ EIV.2020,
              data = regression_df %>% 
                select(fc.per.work.mass.2020,
                       EIV.2020) %>% 
                drop_na())
lm_fc_3 <- lm(fc.per.work.mass.2020 ~ EIV.Calculated,
              data = regression_df %>% 
                select(fc.per.work.mass.2020,
                       EIV.Calculated) %>% 
                drop_na())
lm_fc_3a <- lm(fc.per.work.mass.2020 ~ EIV.Calculated + I(EIV.Calculated^2),
              data = regression_df %>% 
                select(fc.per.work.mass.2020,
                       EIV.Calculated) %>% 
                drop_na())
lm_fc_4 <- lm(fc.per.work.mass.2020 ~ Built.Year,
              data = regression_df %>% 
                select(fc.per.work.mass.2020,
                       Built.Year) %>% 
                drop_na())
lm_fc_5 <- lm(fc.per.work.mass.2020 ~ Owner.Size.fleet,
              data = regression_df %>% 
                select(fc.per.work.mass.2020,
                       Owner.Size.fleet) %>% 
                drop_na())
stargazer(lm_fc_1, lm_fc_2, lm_fc_3, lm_fc_3a, lm_fc_4, lm_fc_5, type = "text")
```
2020
```{r}
regression_df <- fleet_df %>% 
  group_by(Owner) %>% 
  mutate(Owner.Size.fleet = n(),
         Dwt = Dwt/1000) %>% 
  ungroup() %>% 
  select(fc.per.work.mass.2020,
         EIV.Calculated,
         Built.Year,
         Owner.Size.fleet,
         Dwt) %>% 
  drop_na() %>% 
  filter(fc.per.work.mass.2020 <= 20) 

# regression_df %>% 
#   ggplot(aes(Built.Year)) +
#   geom_histogram(binwidth = 1)
# 
# regression_df %>% 
#   ggplot(aes(x = Built.Year,
#              y = fc.per.work.mass.2020)) +
#   geom_point() +
#   geom_smooth(method = 'lm')

EIV_lm <- lm(fc.per.work.mass.2020 ~ EIV.Calculated,
              data = regression_df)
Built_lm <- lm(fc.per.work.mass.2020 ~ Built.Year,
              data = regression_df)
Owner_size_lm <- lm(fc.per.work.mass.2020 ~ Owner.Size.fleet,
              data = regression_df)
Dwt_lm <- lm(fc.per.work.mass.2020 ~ Dwt,
              data = regression_df)
All_lm <- lm(fc.per.work.mass.2020 ~ EIV.Calculated + Built.Year + Owner.Size.fleet + Dwt,
              data = regression_df)
stargazer(EIV_lm, Built_lm, Owner_size_lm, Dwt_lm, All_lm, type = "text")
```

2019
```{r}
regression_df <- fleet_df %>% 
  group_by(Owner) %>% 
  mutate(Owner.Size.fleet = n(),
         Dwt = Dwt/1000) %>% 
  ungroup() %>% 
  select(fc.per.work.mass.2019,
         EIV.Calculated,
         Built.Year,
         Owner.Size.fleet,
         Dwt) %>% 
  drop_na() %>% 
  filter(fc.per.work.mass.2019 <= 20) 

# regression_df %>% 
#   ggplot(aes(Built.Year)) +
#   geom_histogram(binwidth = 1)
# 
# regression_df %>% 
#   ggplot(aes(x = Built.Year,
#              y = fc.per.work.mass.2019)) +
#   geom_point() +
#   geom_smooth(method = 'lm')

EIV_lm <- lm(fc.per.work.mass.2019 ~ EIV.Calculated,
              data = regression_df)
Built_lm <- lm(fc.per.work.mass.2019 ~ Built.Year,
              data = regression_df)
Owner_size_lm <- lm(fc.per.work.mass.2019 ~ Owner.Size.fleet,
              data = regression_df)
Dwt_lm <- lm(fc.per.work.mass.2019 ~ Dwt,
              data = regression_df)
All_lm <- lm(fc.per.work.mass.2019 ~ EIV.Calculated + Built.Year + Owner.Size.fleet + Dwt,
              data = regression_df)
stargazer(EIV_lm, Built_lm, Owner_size_lm, Dwt_lm, All_lm, type = "text")
```

2018
```{r}
regression_df <- fleet_df %>% 
  group_by(Owner) %>% 
  mutate(Owner.Size.fleet = n(),
         Dwt = Dwt/1000) %>% 
  ungroup() %>% 
  select(fc.per.work.mass.2018,
         EIV.Calculated,
         Built.Year,
         Owner.Size.fleet,
         Dwt) %>% 
  drop_na() %>% 
  filter(fc.per.work.mass.2018 <= 20) 

# regression_df %>% 
#   ggplot(aes(Built.Year)) +
#   geom_histogram(binwidth = 1)
# 
# regression_df %>% 
#   ggplot(aes(x = Built.Year,
#              y = fc.per.work.mass.2018)) +
#   geom_point() +
#   geom_smooth(method = 'lm')

EIV_lm <- lm(fc.per.work.mass.2018 ~ EIV.Calculated,
              data = regression_df)
Built_lm <- lm(fc.per.work.mass.2018 ~ Built.Year,
              data = regression_df)
Owner_size_lm <- lm(fc.per.work.mass.2018 ~ Owner.Size.fleet,
              data = regression_df)
Dwt_lm <- lm(fc.per.work.mass.2018 ~ Dwt,
              data = regression_df)
All_lm <- lm(fc.per.work.mass.2018 ~ EIV.Calculated + Built.Year + Owner.Size.fleet + Dwt,
              data = regression_df)
stargazer(EIV_lm, Built_lm, Owner_size_lm, Dwt_lm, All_lm, type = "text")
```


```{r}
regression_df <- fleet_df %>% 
  group_by(Owner) %>% 
  mutate(Owner.Size.fleet = n()) %>% 
  select(fc.per.work.mass.2020, Built.Year, Owner.Size.fleet, EIV.Calculated) %>% 
  drop_na()

lm_fc_1 <- lm(fc.per.work.mass.2020 ~ EIV.Calculated, data = regression_df)
lm_fc_2 <- lm(fc.per.work.mass.2020 ~ Built.Year, data = regression_df)
lm_fc_3 <- lm(fc.per.work.mass.2020 ~ Owner.Size.fleet, data = regression_df)
lm_fc_4 <- lm(fc.per.work.mass.2020 ~ EIV.Calculated + Built.Year + Owner.Size.fleet, data = regression_df)
stargazer(lm_fc_1, lm_fc_2, lm_fc_3, lm_fc_4, type = "text")
```


```{r}
regression_df <- fleet_df %>% 
  group_by(Owner) %>% 
  mutate(Owner.Size.fleet = n()) %>% 
  select(fc.per.work.mass.2020, Built.Year, Owner.Size.fleet, EEDI.2020) %>% 
  drop_na()

lm_fc_1 <- lm(fc.per.work.mass.2020 ~ EEDI.2020, data = regression_df)
lm_fc_2 <- lm(fc.per.work.mass.2020 ~ Built.Year, data = regression_df)
lm_fc_3 <- lm(fc.per.work.mass.2020 ~ Owner.Size.fleet, data = regression_df)
lm_fc_4 <- lm(fc.per.work.mass.2020 ~ EEDI.2020 + Built.Year + Owner.Size.fleet, data = regression_df)
stargazer(lm_fc_1, lm_fc_2, lm_fc_3, lm_fc_4, type = "text")
```

## Efficiency metric vs. fuel consumption
```{r}
fleet_df %>% 
  filter(Fleet.Type == "Bulkers") %>% 
  drop_na(Efficiency.Metric.2020) %>% 
  ggplot(aes(y = FC.Per.Distance.2020,
             x = Technical.Efficiency.2020)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~Efficiency.Metric.2020) +
  ylim(0, 250) +
  xlim(0,40)
```

```{r}
fleet_df %>% 
  drop_na(efficiency.metric.2020) %>% 
  ggplot(aes(y = fc.per.work.mass.2020,
             x = technical.efficiency.2020)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~efficiency.metric.2020) +
  ylim(0, 250) +
  xlim(0,54)
```

```{r}
fleet_df %>% 
  drop_na(efficiency.metric.2020) %>% 
  ggplot(aes(y = total.fc.2020,
             x = technical.efficiency.2020)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~efficiency.metric.2020) +
  # ylim(0, 250) +
  xlim(0,40)
```

## Design efficiency
```{r}
fleet_df %>% 
  ggplot(aes(x = Main.Engine.Power.kW, fill = Size.Category)) +
  geom_histogram(binwidth = 500)
```

```{r}
fleet_df %>% 
  drop_na(design.efficiency) %>% 
  ggplot(aes(x = Dwt,
             y = design.efficiency,
             shape = Size.Category,
             group = Size.Category,
             color = Built.Date)) +
  geom_point() +
  # xlim(2, 10) +
  ylim(0, 1.5) +
  geom_smooth(method = 'lm', color = 'black')
```

```{r}
fleet_df %>% 
  pivot_longer(contains("EEDI.2"),
               values_to = "EEDI",
               names_to = "reporting.period",
               names_prefix = "EEDI.") %>% 
  drop_na(EEDI) %>% 
  group_by(CVN) %>% 
  summarise(EEDI = min(EEDI, na.rm = TRUE),
            Dwt = first(Dwt),
            Size.Category = first(Size.Category),
            Built.Date = first(Built.Date)) %>% 
  ggplot(aes(x = Dwt,
             y = EEDI,
             shape = Size.Category,
             group = Size.Category,
             color = Built.Date)) +
  geom_point() +
  # xlim(2, 10) +
  ylim(0, 20) +
  geom_smooth(method = 'lm', color = 'black')
```

```{r}
fleet_df %>% 
  # filter(Size.Category == "Panamax") %>% 
  pivot_longer(contains("EIV.2"),
               values_to = "EIV",
               names_to = "reporting.period",
               names_prefix = "EIV.") %>% 
  drop_na(EIV) %>% 
  group_by(CVN) %>% 
  summarise(EIV = min(EIV, na.rm = TRUE),
            Dwt = first(Dwt),
            Size.Category = first(Size.Category),
            Built.Date = first(Built.Date)) %>% 
  ggplot(aes(x = Dwt,
             y = EIV,
             shape = Size.Category,
             group = Size.Category,
             color = Built.Date)) +
  geom_point() +
  # xlim(0, 110E3) +
  ylim(0, 10) +
  geom_smooth(method = 'lm', color = 'black')
```

```{r}
fleet_df %>% 
  ggplot(aes(x = Dwt,
             y = EIV.Calculated,
             shape = Size.Category,
             group = Size.Category,
             color = Built.Date)) +
  geom_point() +
  # xlim(0, 110E3) +
  ylim(0, 10) +
  geom_smooth(method = 'lm', color = 'black')
```

```{r}
fleet_df %>% 
  select(-contains("fc.per.work.mass.o")) %>% 
  pivot_longer(contains("fc.per.work.mass."),
               values_to = "fc.per.work.mass",
               names_to = "reporting.period",
               names_prefix = "fc.per.work.mass.") %>% 
  drop_na(fc.per.work.mass) %>%
  group_by(CVN) %>%
  summarise(fc.per.work.mass = mean(fc.per.work.mass, na.rm = TRUE),
            Dwt = first(Dwt),
            Size.Category = first(Size.Category),
            Built.Date = first(Built.Date)) %>%
  ggplot(aes(x = Dwt,
             y = fc.per.work.mass,
             shape = Size.Category,
             group = Size.Category,
             color = Built.Date)) +
  geom_point() +
  # xlim(2, 10) +
  ylim(0, 1.5) +
  geom_smooth(method = 'lm', color = 'black')
```

```{r}
fleet_df %>% 
  pivot_longer(contains("EEDI.2"),
               values_to = "EEDI",
               names_to = "reporting.period",
               names_prefix = "EEDI.") %>% 
  drop_na(EEDI, design.efficiency) %>% 
  group_by(CVN) %>% 
  summarise(EEDI = min(EEDI, na.rm = TRUE),
            design.efficiency = min(design.efficiency, na.rm = TRUE),
            Size.Category = first(Size.Category)) %>% 
  ggplot(aes(x = EEDI,
             y = design.efficiency,
             color = Size.Category)) +
  geom_point() +
  xlim(2, 10) +
  ylim(0, 1) +
  geom_smooth(method = 'lm') +
  facet_wrap(~Size.Category)
```
These don't fall along the regression line as well as I would have hoped.

```{r}
fleet_df %>% 
  pivot_longer(contains("EIV.2"),
               values_to = "EIV",
               names_to = "reporting.period",
               names_prefix = "EIV.") %>% 
  drop_na(EIV, design.efficiency) %>% 
  group_by(CVN) %>% 
  summarise(EIV = min(EIV, na.rm = TRUE),
            design.efficiency = min(design.efficiency, na.rm = TRUE),
            Size.Category = first(Size.Category)) %>% 
  ggplot(aes(x = EIV,
             y = design.efficiency,
             color = Size.Category)) +
  geom_point() +
  xlim(2, 10) +
  ylim(0, 1) +
  geom_smooth(method = 'lm') +
  facet_wrap(~Size.Category)
```
This is better. Need to check definition of EIV.


### EIV Calculated
```{r}
fleet_df %>% 
  pivot_longer(contains("EIV.2"),
               values_to = "EIV",
               names_to = "reporting.period",
               names_prefix = "EIV.") %>% 
  drop_na(EIV, EIV.Calculated) %>% 
  group_by(CVN) %>% 
  summarise(EIV = min(EIV, na.rm = TRUE),
            EIV.Calculated = first(EIV.Calculated),
            Size.Category = first(Size.Category)) %>% 
  ggplot(aes(x = EIV,
             y = EIV.Calculated,
             color = Size.Category)) +
  geom_point() +
  xlim(0, 30) +
  ylim(0, 30) +
  geom_smooth(method = 'lm') +
  facet_wrap(~Size.Category)
```

```{r}
fleet_df %>% 
  pivot_longer(contains("EIV.2"),
               values_to = "EIV",
               names_to = "reporting.period",
               names_prefix = "EIV.") %>% 
  drop_na(EIV, EIV.Main.Calculated) %>% 
  group_by(CVN) %>% 
  summarise(EIV = min(EIV, na.rm = TRUE),
            EIV.Main.Calculated = first(EIV.Main.Calculated),
            Size.Category = first(Size.Category)) %>% 
  ggplot(aes(x = EIV,
             y = EIV.Main.Calculated,
             color = Size.Category)) +
  geom_point() +
  xlim(0, 30) +
  ylim(0, 30) +
  geom_smooth(method = 'lm') +
  facet_wrap(~Size.Category)
```

```{r}
fleet_df %>% 
  ggplot(aes(x = Built.Date,
             y = design.efficiency,
             color = Size.Category)) +
  geom_point() +
  facet_wrap(~Size.Category, nrow = 1)
```

# Size
```{r}
fleet_df %>% 
  pivot_longer(c(Beam.Mid..m,
                 Beam.Ext..m.,
                 Draught..m.,
                 Depth.Moulded..m.,
                 Air.Draft.From.Keel,
                 Air.Draft.From.Water.Line..m.)) %>% 
  ggplot(aes(x = value,
             fill = Size.Category)) +
  geom_histogram(position = 'identity', 
                 alpha = 0.6, 
                 binwidth = 1) +
  facet_wrap(~name, scales = 'free')
```
