---
title: "Shipping Emissions"
output: html_document
---
```{r}
library(tidyverse)
library(reticulate)
use_python("/usr/bin/python3")
```

# Data Formatting
Monitoring Reporting Verification EU Emissions reports
```{r, child = 'MRV_Data.rmd', eval = TRUE}
```

World Fleet Register
```{r, child = 'WFR_Data.rmd', eval = FALSE}
```

## AIS
```{r, eval = FALSE}
py_run_file('AIS_PreData_Ingest.py')
```

MMSI lists by ship type for subsetting AIS data:
```{r eval = FALSE}
load("data/bulkers_WFR.Rda")
bulkers_mmsi <- bulkers_df %>% 
  distinct(MMSI) %>% 
  drop_na() %>% 
  write_csv("data/bulkers_mmsi.csv")

load("data/containerships_WFR.Rda")
containerships_mmsi <- containerships_df %>% 
  distinct(MMSI) %>% 
  drop_na() %>% 
  write_csv("data/containerships_mmsi.csv")
```

```{r, eval = FALSE}
py_run_file('AIS_PreData_Subset.py')
```

High memory requirements, use Compute Canada
```{r, eval = FALSE}
py_run_file('AIS_Data_Index.py')
```

High memory requirements, use Compute Canada
```{r, eval = FALSE}
py_run_file('AIS_Data_Sort.py')
```

```{r, eval = FALSE}
py_run_file('AIS_Data_IDs.py')
```

```{r, child = 'AIS_Data_IDs.rmd', eval = FALSE}
```

```{r}
py_run_file('AIS_Data_Static.py')
```

```{r, eval = FALSE}
py_run_file('AIS_Data_Combine.py')
```

# Fixed ship component of power consumption
```{r, eval = FALSE}
py_run_file('WFR Calculations_Fixed_Power.py')
```


# AIS Calculations
```{r, eval = FALSE}
py_run_file('EEZ_EU_Assignment.py')
```

```{r, eval = FALSE}
py_run_file('AIS_Calculations.py')
```
Files useful for troubleshooting and verifying AIS_calculations:
- AIS_Calculations_Troubleshooting.py
- AIS_Calculations_JumpTesting.py
- EU_portcall_verification.qgz


```{r, eval = FALSE}
py_run_file('AIS_Calculations_Interp.py')
```

```{r, eval = FALSE}
py_run_file('AIS_Calculations_Interp_Stats.py')
```

```{r, child = 'AIS_Calculations_Interp_Stats_Explore.qmd', eval = FALSE}
```


```{r, eval = FALSE}
py_run_file('AIS_Calculations_PotentialPortcalls.py')
```

```{r, eval = FALSE}
py_run_file('AIS_Calculations_PP_Coast_EU.py')
```

# Fuel Consumption & Yearly Aggregation

```{r, eval = FALSE}
py_run_file('EU_yearly_agg.py')
```





