---
title: "week3"
output: html_document
---

```{r}
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(stargazer)
```

```{r}
load("data/All_join.Rda")

fleet_df$Owner.Size <- factor(fleet_df$Owner.Size,
                                c("Very Small (1-5)",
                                    "Small (6-10)",
                                    "Medium (11-20)",
                                    "Large (21-50)",
                                    "Very Large (51-100)",
                                    "Extra Large (100+)"))

```

#### 2020 emission factor distribution
```{r}
emit <- na.omit(c(fleet_df$MRV.Emission.Factor.2020
                  #, fleet_df$MRV.Emission.Factor.2019
                  #, fleet_df$MRV.Emission.Factor.2018
                  ))
upper <- quantile(x = emit,probs = .99)
lower <- quantile(x = emit,probs = .01)
emit_filtered <- emit[emit < upper & emit > lower]
hist(emit_filtered, main = paste("Histogram of 2020 MRV emission factor"))



```



#### stacked 2020 emission factor diagram
```{r}

temp_df <- fleet_df %>% filter(MRV.Emission.Factor.2020 < quantile(x = MRV.Emission.Factor.2020,probs = .99, na.rm = TRUE) &
                                 MRV.Emission.Factor.2020 > quantile(x = MRV.Emission.Factor.2020 ,probs = .01, na.rm = TRUE))


temp_df %>% ggplot(aes(x = MRV.Emission.Factor.2020, fill = SOx.Scrubber.Indicator))+
  geom_histogram()

```





#### fc per work mass vs co2 per work mass
```{r}
temp_df <- fleet_df

temp_df$co2perworkmass2020 <- temp_df$MRV.FC.Per.Work.Mass.2020 * temp_df$MRV.Emission.Factor.2020
temp_df$co2perworkmass2019 <- temp_df$MRV.FC.Per.Work.Mass.2019 * temp_df$MRV.Emission.Factor.2019
temp_df$co2perworkmass2018 <- temp_df$MRV.FC.Per.Work.Mass.2018 * temp_df$MRV.Emission.Factor.2018

temp_df %>% 
  #select(-contains("fc.per.distance")) %>% 
  pivot_longer(contains("co2perworkmass"), values_to = "co2perworkmass", names_to = "reporting.temp") %>%
  pivot_longer(contains("fc.per.work.mass"), values_to = "fcperworkmass", names_to ="reporting.period") %>% 
  filter(co2perworkmass < quantile(x = co2perworkmass, probs = 0.99, na.rm = TRUE) &
           co2perworkmass > quantile(x = co2perworkmass, probs = 0.01, na.rm = TRUE) &
           fcperworkmass < quantile(x = fcperworkmass, probs = 0.99, na.rm = TRUE) &
           fcperworkmass > quantile(x = fcperworkmass, probs = 0.01, na.rm = TRUE)) %>% # filter outliers
  ggplot(aes(y = co2perworkmass, 
             x = fcperworkmass,
             color = reporting.temp)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  # coord_cartesian(ylim = c(0, 100)) +
  #ylim(0,20) +
  facet_wrap(~Size.Category)



```


#### CO2 per work mass vs built year
```{r}
temp_df <- fleet_df

temp_df$co2perworkmass2020 <- temp_df$MRV.FC.Per.Work.Mass.2020 * temp_df$MRV.Emission.Factor.2020
temp_df$co2perworkmass2019 <- temp_df$MRV.FC.Per.Work.Mass.2019 * temp_df$MRV.Emission.Factor.2019
temp_df$co2perworkmass2018 <- temp_df$MRV.FC.Per.Work.Mass.2018 * temp_df$MRV.Emission.Factor.2018

temp_df %>% 
  #select(-contains("fc.per.distance")) %>% 
  pivot_longer(contains("co2perworkmass"), values_to = "co2perworkmass", names_to = "reporting.temp") %>%
  pivot_longer(contains("fc.per.work.mass"), values_to = "fcperworkmass", names_to ="reporting.period") %>% 
  filter(co2perworkmass < quantile(x = co2perworkmass, probs = 0.99, na.rm = TRUE) &
           co2perworkmass > quantile(x = co2perworkmass, probs = 0.01, na.rm = TRUE) &
           fcperworkmass < quantile(x = fcperworkmass, probs = 0.99, na.rm = TRUE) &
           fcperworkmass > quantile(x = fcperworkmass, probs = 0.01, na.rm = TRUE)) %>% # filter outliers
  ggplot(aes(y = co2perworkmass, 
             x = Built.Year,
             color = reporting.temp)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  # coord_cartesian(ylim = c(0, 100)) +
  #ylim(0,20) +
  facet_wrap(~Size.Category)
```

##### MRV dist vs AIS dist
```{r}
temp_df <- fleet_df

temp_df %>% 
  #select(-contains("fc.per.distance")) %>% 
  pivot_longer(contains("mrv.eu.distance"), values_to = "mrvdist", names_to = "reporting.temp") %>%
  pivot_longer(contains("ais.distance"), values_to = "aisdist", names_to ="reporting.period") %>% 
  filter(mrvdist < quantile(x = mrvdist, probs = 0.99, na.rm = TRUE) &
           mrvdist > quantile(x = mrvdist, probs = 0.01, na.rm = TRUE) &
           aisdist < quantile(x = aisdist, probs = 0.99, na.rm = TRUE) &
           aisdist > quantile(x = aisdist, probs = 0.01, na.rm = TRUE)) %>% # filter outliers
  ggplot(aes(y = mrvdist, 
             x = aisdist,
             color = reporting.temp)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  # coord_cartesian(ylim = c(0, 100)) +
  #ylim(0,20) +
  facet_wrap(~Size.Category)


```


#### ratio of distance travelled in mrv vs ais
```{r}
temp_df <- fleet_df

temp_df$ratio2020 <- temp_df$AIS.Distance.2020 / temp_df$MRV.EU.Distance.2020 
temp_df$ratio2019 <- temp_df$AIS.Distance.2019 / temp_df$MRV.EU.Distance.2019

ratio <- na.omit(c(temp_df$ratio2020, temp_df$ratio2019))

upper <- quantile(x = ratio,probs = .99)
lower <- quantile(x = ratio,probs = .01)
ratio_filtered <- ratio[ratio < upper & ratio > lower]

hist(ratio_filtered, breaks = 40, main = paste("Histogram of AIS dist / MRV dist for 2019, 2020"))


```


```{r}
temp_df <- fleet_df

temp_df$ratio2020 <- temp_df$AIS.Distance.2020 / temp_df$MRV.EU.Distance.2020 
temp_df$ratio2019 <- temp_df$AIS.Distance.2019 / temp_df$MRV.EU.Distance.2019

temp_df$ratiochange <- temp_df$ratio2020 - temp_df$ratio2019

ratio <- na.omit(c(temp_df$ratiochange))

upper <- quantile(x = ratio,probs = .99)
lower <- quantile(x = ratio,probs = .01)
ratio_change <- ratio[ratio < upper & ratio > lower]

hist(ratio_change, breaks = 40, main = paste("Histogram of change of AIS dist / MRV dist from 2019 to 2020"))


```


##### total co2 scatter
```{r}
temp_df <- fleet_df

c2018 <- sum(temp_df$MRV.Total.CO2.2018, na.rm = TRUE)

c2019 <- sum(temp_df$MRV.Total.CO2.2019, na.rm = TRUE)

c2020 <- sum(temp_df$MRV.Total.CO2.2020, na.rm = TRUE)

plot(y = c(c2018,c2019,c2020), x = c(0,1,2))




```

##### total dist scatter
```{r}
temp_df <- fleet_df

c2018 <- sum(temp_df$MRV.EU.Distance.2018, na.rm = TRUE)

c2019 <- sum(temp_df$MRV.EU.Distance.2019, na.rm = TRUE)

c2020 <- sum(temp_df$MRV.EU.Distance.2020, na.rm = TRUE)

plot(y = c(c2018,c2019,c2020), x = c(0,1,2))




```
##### total dist scatter filtered by scrubber
```{r}
temp_df <- fleet_df %>% filter(SOx.Scrubber.Indicator == "Y")

c2018 <- sum(temp_df$MRV.EU.Distance.2018, na.rm = TRUE)

c2019 <- sum(temp_df$MRV.EU.Distance.2019, na.rm = TRUE)

c2020 <- sum(temp_df$MRV.EU.Distance.2020, na.rm = TRUE)

plot(y = c(c2018,c2019,c2020), x = c(0,1,2))




```


##### total dist scatter filtered by scrubber
```{r}
temp_df <- fleet_df %>% filter(SOx.Scrubber.Indicator == "Y")

c2018 <- sum(temp_df$MRV.Total.CO2.2018, na.rm = TRUE)

c2019 <- sum(temp_df$MRV.Total.CO2.2019, na.rm = TRUE)

c2020 <- sum(temp_df$MRV.Total.CO2.2020, na.rm = TRUE)

plot(y = c(c2018,c2019,c2020), x = c(0,1,2))




```
##### total dist scatter filtered by no scrubber
```{r}
temp_df <- fleet_df %>% filter(SOx.Scrubber.Indicator != "Y")

c2018 <- sum(temp_df$MRV.EU.Distance.2018, na.rm = TRUE)

c2019 <- sum(temp_df$MRV.EU.Distance.2019, na.rm = TRUE)

c2020 <- sum(temp_df$MRV.EU.Distance.2020, na.rm = TRUE)

plot(y = c(c2018,c2019,c2020), x = c(0,1,2))




```


##### total dist scatter filtered by no scrubber
```{r}
temp_df <- fleet_df %>% filter(SOx.Scrubber.Indicator != "Y")

c2018 <- sum(temp_df$MRV.Total.CO2.2018, na.rm = TRUE)

c2019 <- sum(temp_df$MRV.Total.CO2.2019, na.rm = TRUE)

c2020 <- sum(temp_df$MRV.Total.CO2.2020, na.rm = TRUE)

plot(y = c(c2018,c2019,c2020), x = c(0,1,2))




```
